% $Id: introduction.tex 142 2012-12-22 10:41:32Z danbos $
\chapter{Introduction}
\label{ch:introduction}
\noindent
This project is a very basic prototype for a project that will be open source: A community tool for role players. The idea for it came out of a recent issue where one person's code was hosted by that person for several different role playing chats, and when he no longer had the time or energy to deal with the hosting, he forced the chats to close, as he was not interested in sharing it.

The current prototype only includes the beginnings of a user system and the ability to post on a communal wall. That communal wall is going to be expanded to several different kinds of messages, and the user system will become more full-fledged, with individualised profiles and e-mail reminders. One of the key requirements to the future implementations is the ability to easily export/import all of the data from one installation to another, as well as be open sourced so that others can pick it up if necessary.

\section{Disposition}
\label{ch:disposition}
\noindent
Chapter 1 describes the reasons behind the project and what is intended to be accomplished.

Chapter 2 describes separation of concerns in modern web development, followed by oWasp top-10 list of vulnerabilities for 2013, with several examples of the exploits in the wild.

Chapter 3 describes the technologies going into building the project, and how the different vulnerabilities are addressed, when relevant.

Chapter 4 describes the final discussions on how to think when starting a project security-first.


